Robust Clustering for Ad Hoc Cognitive Radio Network

### 2017-04-17 journal-draft/ettdoc.tex
change 'all its neighbors' to all of its neighbors

\item \textit{Neighborhood connectivity degree} $g_i$: the number of CCs which are available for $i$ and all its neighbors.
$g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$, which represents the ability of $i$ to form a robust cluster with its neighbors.
\end{itemize}
Individual connectivity degree $d_i$ and neighborhood connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where every node's connectivity vector is shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph of the example CRN and the connectivity vector $(d_i, g_i)$ for each node. The desired cluster size $\delta =3$. The sets of the indices of the available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range.}
	\label{fig1}
\end{figure}

\subsubsection{Determining Cluster Heads and Forming Clusters}
The procedure of determining the cluster heads is as follows.
Each CR node decides whether it is a cluster head by comparing its connectivity vector with its neighbors.
When CR node $i$ has lower individual connectivity degree than all its neighbors except for those which have already identified to be cluster heads, node $i$ becomes a cluster head.
If there is another CR node $j$ in its neighborhood, which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ and $d_j < d_{k}, \forall k\in \text{Nb}(j)\setminus \{\Lambda\cup i\}$ where $\Lambda$ denotes the cluster heads, then the node between $i$ and $j$, which has higher neighborhood connectivity degree will become the cluster head.
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes the cluster head.
%
The node which is identified as a cluster head broadcasts a message to notify its neighbors of this change, and its neighbors which are not cluster heads become cluster members
\footnote{The reason for the occurrence of the cluster heads in the neighborhood of a new cluster head will be explained in Section~\ref{ross_p1_guarantee_ccc} and \ref{ross_p2_cluster_pruning})}.
The pseudo code for the cluster head decision and the initial cluster formation is shown in Algorithm~\ref{alg0} in the appendix.

After receiving the notification from a cluster head, a CR node $i$ is aware that it becomes a member of a cluster. 
Consequently, $i$ sets its individual connectivity degree to a positive number $M > |\mathcal{K}| \cdot N$, and broadcasts the new individual connectivity degree to all its neighbors. 
%wrong:
%Note that there is no cluster head which receives the notification on cluster head eligibility, and there is a Lemma on this.
%\begin{theorem}{lemma}
%There is no cluster head receiving the notification message on cluster head eligibility, which is sent from a node which newly becomes a cluster head.
%\end{theorem}
%\begin{proof}
%We use contradiction to prove. 
%Assume a CR node $j$ is cluster head, and receives a notification on cluster head eligibility form CR node $i$. 
%According to the assumption on the reciprocal communication link, $i$ should have received the notification message from $j$ when $j$ becomes cluster head.
%Then CR node $i$ will set its individual connectivity degree to zero, and has no possibility to become a cluster head.
%\end{proof}
When a CR node $i$ is associated to multiple clusters, \ie $i$ has received multiple notifications from different cluster heads, $d_i$ is still set to be $M$. 
%In other words, with this manipulation on its individual connectivity degree, every CR node in the network will be involved in the clustering process, the proof will be given later.
The manipulation of the individual connectivity degree of the cluster members accelerates the decision on the cluster heads.
%, which enables the nodes which are outside of the cluster can possibly become cluster heads or can also be included into other clusters.
%We have the following theorem to show that when there is no one secondary which shares no CC with any other secondary users, \ie
%all secondary users' individual connectivity degrees are greater than zero, then every secondary user will eventually be either integrated into a certain cluster, or becomes a cluster head.
We have the following theorem to show that every secondary user will eventually be either integrated into a certain cluster or becomes a cluster head.
\begin{theorem}
\label{clustering:theorem}
Given a CRN, it takes at most $N$ steps that every secondary user either becomes cluster head, or gets included into at least one cluster.
\end{theorem}
Here, by \textit{step} we mean one secondary user executing Algorithm~\ref{clustering:theorem} for one time.
The Proof is in Appendix~\ref{proof_clustering:theorem}.
%
The procedure of the proof also illustrates the maximal time needed to conduct Algorithm~\ref{clustering:theorem}. 
Consider an extreme scenario, where all the secondary nodes sequentially execute Algorithm~\ref{alg0}, \ie they constitute a list as discussed in the example in the proof.
%
If one step can be finished within a certain time span $T$, then the total time needed for the network to conduct Algorithm~\ref{clustering:theorem} is $N*T$.
As Algorithm~\ref{alg0} can be executed concurrently by different secondary users, the needed time can be considerably reduced.
%According to Theorem~\ref{clustering:theorem}, Phase I completes within a reasonable amount of time.
%
%This judgement is conducted periodically, and phase I ends after every node ascertains it is cluster head or not.
If we apply Algorithm~\ref{alg0} to the example shown in Figure~\ref{fig1}, then the outcome is shown in Figure~\ref{fig2}.
Node $B$ and $H$ have the same individual connectivity degree, i.e., $d_B=d_H$. As $g_H=2>g_B=1$, node $H$ becomes the cluster head and cluster $C(H)$ is $\{H, B, A, G\}$.

\subsubsection{The Existence of Common Channels}
\label{ross_p1_guarantee_ccc}
%After deciding itself being cluster head, CR node broadcasts to notify its neighbours on the dedicated control channel, meanwhile, $i$'s initial cluster is formed immediately, which is $i$'s neighbourhood except for those nodes which have become cluster heads, \ie $C_i=(\text{Nb}_i\setminus \text{CHs})\cup i$.
After executing Algorithm~\ref{alg0}, certain formed clusters may not possess any CCs.
As decreasing cluster size increases the CCs within a cluster, for those clusters having no CCs, certain nodes need to be eliminated to obtain at least one CC.
The sequence of elimination is performed according to an ascending list of nodes which are sorted by the number of common channels between the nodes and the cluster head. 
In other words, the cluster member which has the least common channels with the cluster head is excluded first.
If there are multiple nodes having the same number of common channels with the cluster head, the node whose elimination brings in more common channels will be excluded.
If this criterion meets a tie, the tie will be broken by deleting the node with smaller node ID.
It is possible that the cluster head excludes all its neighbors, resulting in a singleton cluster which is composed by itself.
%At the end of this procedure every formed cluster has at least one common channel.
%The pseudo code for cluster head to obtain at least one common channel is shown in Algorithm~\ref{alg_size_control_available_CCC}.
The pseudo code for this procedure is shown in Algorithm~\ref{alg_size_control_available_CCC}.
As to the nodes which are eliminated from the previous clusters, they restore their original individual connectivity degrees, then execute Algorithm~\ref{alg0} and become either cluster heads or get included into other clusters afterwards according to Theorem~\ref{clustering:theorem}.

>>>>

\item \textit{Neighborhood connectivity degree} $g_i$: the number of CCs which are available for $i$ and all of its neighbors.
$g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$, which represents the ability of $i$ to form a robust cluster with its neighbors.
\end{itemize}
Individual connectivity degree $d_i$ and neighborhood connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where every node's connectivity vector is shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph of the example CRN and the connectivity vector $(d_i, g_i)$ for each node. The desired cluster size $\delta =3$. The sets of the indices of the available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range.}
	\label{fig1}
\end{figure}

\subsubsection{Determining Cluster Heads and Forming Clusters}
The procedure of determining the cluster heads is as follows.
Each CR node decides whether it is a cluster head by comparing its connectivity vector with its neighbors.
When CR node $i$ has lower individual connectivity degree than all of its neighbors except for those which have already identified to be cluster heads, node $i$ becomes a cluster head.
If there is another CR node $j$ in its neighborhood, which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ and $d_j < d_{k}, \forall k\in \text{Nb}(j)\setminus \{\Lambda\cup i\}$ where $\Lambda$ denotes the cluster heads, then the node between $i$ and $j$, which has higher neighborhood connectivity degree will become the cluster head.
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes the cluster head.
%
The node which is identified as a cluster head broadcasts a message to notify its neighbors of this change, and its neighbors which are not cluster heads become cluster members
\footnote{The reason for the occurrence of the cluster heads in the neighborhood of a new cluster head will be explained in Section~\ref{ross_p1_guarantee_ccc} and \ref{ross_p2_cluster_pruning})}.
The pseudo code for the cluster head decision and the initial cluster formation is shown in Algorithm~\ref{alg0} in the appendix.

After receiving the notification from a cluster head, a CR node $i$ is aware that it becomes a member of a cluster. 
Consequently, $i$ sets its individual connectivity degree to a positive number $M > |\mathcal{K}| \cdot N$, and broadcasts the new individual connectivity degree to all of its neighbors. 
%wrong:
%Note that there is no cluster head which receives the notification on cluster head eligibility, and there is a Lemma on this.
%\begin{theorem}{lemma}
%There is no cluster head receiving the notification message on cluster head eligibility, which is sent from a node which newly becomes a cluster head.
%\end{theorem}
%\begin{proof}
%We use contradiction to prove. 
%Assume a CR node $j$ is cluster head, and receives a notification on cluster head eligibility form CR node $i$. 
%According to the assumption on the reciprocal communication link, $i$ should have received the notification message from $j$ when $j$ becomes cluster head.
%Then CR node $i$ will set its individual connectivity degree to zero, and has no possibility to become a cluster head.
%\end{proof}
When a CR node $i$ is associated to multiple clusters, \ie $i$ has received multiple notifications from different cluster heads, $d_i$ is still set to be $M$. 
%In other words, with this manipulation on its individual connectivity degree, every CR node in the network will be involved in the clustering process, the proof will be given later.
The manipulation of the individual connectivity degree of the cluster members accelerates the decision on the cluster heads.
%, which enables the nodes which are outside of the cluster can possibly become cluster heads or can also be included into other clusters.
%We have the following theorem to show that when there is no one secondary which shares no CC with any other secondary users, \ie
%all secondary users' individual connectivity degrees are greater than zero, then every secondary user will eventually be either integrated into a certain cluster, or becomes a cluster head.
We have the following theorem to show that every secondary user will eventually be either integrated into a certain cluster or becomes a cluster head.
\begin{theorem}
\label{clustering:theorem}
Given a CRN, it takes at most $N$ steps that every secondary user either becomes cluster head, or gets included into at least one cluster.
\end{theorem}
Here, by \textit{step} we mean one secondary user executing Algorithm~\ref{clustering:theorem} for one time.
The Proof is in Appendix~\ref{proof_clustering:theorem}.
%
The procedure of the proof also illustrates the maximal time needed to conduct Algorithm~\ref{clustering:theorem}. 
Consider an extreme scenario, where all the secondary nodes sequentially execute Algorithm~\ref{alg0}, \ie they constitute a list as discussed in the example in the proof.
%
If one step can be finished within a certain time span $T$, then the total time needed for the network to conduct Algorithm~\ref{clustering:theorem} is $N*T$.
As Algorithm~\ref{alg0} can be executed concurrently by different secondary users, the needed time can be considerably reduced.
%According to Theorem~\ref{clustering:theorem}, Phase I completes within a reasonable amount of time.
%
%This judgement is conducted periodically, and phase I ends after every node ascertains it is cluster head or not.
If we apply Algorithm~\ref{alg0} to the example shown in Figure~\ref{fig1}, then the outcome is shown in Figure~\ref{fig2}.
Node $B$ and $H$ have the same individual connectivity degree, i.e., $d_B=d_H$. As $g_H=2>g_B=1$, node $H$ becomes the cluster head and cluster $C(H)$ is $\{H, B, A, G\}$.

\subsubsection{The Existence of Common Channels}
\label{ross_p1_guarantee_ccc}
%After deciding itself being cluster head, CR node broadcasts to notify its neighbours on the dedicated control channel, meanwhile, $i$'s initial cluster is formed immediately, which is $i$'s neighbourhood except for those nodes which have become cluster heads, \ie $C_i=(\text{Nb}_i\setminus \text{CHs})\cup i$.
After executing Algorithm~\ref{alg0}, certain formed clusters may not possess any CCs.
As decreasing cluster size increases the CCs within a cluster, for those clusters having no CCs, certain nodes need to be eliminated to obtain at least one CC.
The sequence of elimination is performed according to an ascending list of nodes which are sorted by the number of common channels between the nodes and the cluster head. 
In other words, the cluster member which has the least common channels with the cluster head is excluded first.
If there are multiple nodes having the same number of common channels with the cluster head, the node whose elimination brings in more common channels will be excluded.
If this criterion meets a tie, the tie will be broken by deleting the node with smaller node ID.
It is possible that the cluster head excludes all of its neighbors, resulting in a singleton cluster which is composed by itself.
%At the end of this procedure every formed cluster has at least one common channel.
%The pseudo code for cluster head to obtain at least one common channel is shown in Algorithm~\ref{alg_size_control_available_CCC}.
The pseudo code for this procedure is shown in Algorithm~\ref{alg_size_control_available_CCC}.
As to the nodes which are eliminated from the previous clusters, they restore their original individual connectivity degrees, then execute Algorithm~\ref{alg0} and become either cluster heads or get included into other clusters afterwards according to Theorem~\ref{clustering:theorem}.
